Coastal Flooding (surface water, highway, sewer and watercourse flooding) across Torbay, UK

### 2016-11-17 EU-Circle/eu-circle_case-study-3.Rmd
EU-Circle case study 3

```{r precipitation-intensity, warning=FALSE, message=FALSE}
## Show the mean seasonal cycle in the wet-day mean precipitation
plot(MU)
```

@@

```{r wet-day frequency, warning=FALSE, message=FALSE}
## Show the mean seasonal cycle in the wet-day frequency
plot(FW)
```

@@

```{r sesonal-precipitation, warning=FALSE, message=FALSE}
## Show the mean seasonal cycle in the total precipitation
MP <- 30*MP
attr(MP,'unit') <- 'mm/month'
plot(MP)
```

@@

```{r winter-rain, warning=FALSE, message=FALSE}
## Estimate the PCA for the annual cycle to identify the general features:
pca.MU <- PCA(MU,n=10)
plot(pca.MU)
```

@@

@@

```{r}
plot(gloss)
```

@@

```{r}
plot(newlyn)
```

@@

```{r sonel, warning=FALSE, message=FALSE}
sonel <- aggregate(station.sonel(),year,FUN='max',na.rm=TRUE)
plot(sonel)
```

@@

```{r trend-map, warning=FALSE, message=FALSE}
## Fudge it - pick some windstations, but do not use their data. Reset the
## coordinates to select a triangle of SLP from the EOBS data.
ss <- select.station(param='slp',src='ecad',cntr=c('France','UNITED KINGDOM'))
fx <- station(ss)
triangle <- subset(fx,is=1:3)
## The stations are not situated quite where there is valid slp data in the EOBS dataset
## It's only their coordinates that is used for interpolating SLP for three stations
## Using station objects as is-object makes the results a station object
attr(triangle,'longitude') <- c(-4,-2,-1.25)
attr(triangle,'latitude') <- c(50.50,51,49)
pp.eobs <- regrid(slp.eobs,is=triangle)
## Use triagulation for stations to estimate the wind
uv <- geostrophicwind(pp.eobs)
plot(annual(uv,nmin=300),main='zonal and meridional wind components',
     sub='Estimated from triangulation of mean sea-level pressure from EOBS',zoom=7)
```

@@

```{r trend-event-2006, warning=FALSE, message=FALSE}
plot(annual(uv,nmin=300),main='zonal and meridional wind components',
     xlim=c(2000,2010))
```

@@

```{r zonal-wind-check1, warning=FALSE, message=FALSE}
## Trend in upper wind direction
uam <- annual(subset(uv,is=2),nmin=300)
ds.u1 <- DS(uam,eof.slp,eofs=1:20)
plot(ds.u1)
```

@@

```{r zonal-wind-check2, warning=FALSE, message=FALSE}
## Trend in upper wind direction
ds.u2 <- DS(subset(uam,it=c(1950,2005)),eof.slp,eofs=1:20)
plot(ds.u2)
```

@@

```{r reanalysis-wind-speed, warning=FALSE, message=FALSE}
## 
UV <- geostrophicwind(SLP)
```

@@

```{r EOFs-in-precip-mu, warning=FALSE, message=FALSE}
## The area mean wet-day mean precipitation
eof.mu <- EOF(mu)
## The area mean wet-day frequency
plot(eof.mu)

```{r EOFs-in-precip-fw, warning=FALSE, message=FALSE}
## Show maps of trends for the wet-day mean precipitation
eof.fw <- EOF(fw)
## Map of trends in wet-day frequency:
plot(eof.fw)
```

```{r cca, warning=FALSE, message=FALSE}
## The wet-day frequency and SLP
cca.fw <- CCA(eof.fw,eof.slp)
plot(cca.fw)
cca.mu <- CCA(eof.mu,eof.es)
plot(cca.mu)
```

```{r DS, warning=FALSE, message=FALSE}
## The wet-day frequency and SLP
ds.fw <- DS(subset(eof.fw,ip=1:4),eof.slp,eofs=1:5)
plot(ds.fw)
ds.mu <- DS(subset(eof.mu,ip=1:4),eof.es,eofs=1:7)
plot(ds.mu)
```

Extract the residual and examine to see if there is any temporal-stpatial structure left: a tell-tale sign for noise is a flat eigenspectrum

```{r residual-for-mu, warning=FALSE, message=FALSE}
## The residual of the wet-day precipitation
z.mu <- as.field(ds.mu)
#res.mu <- as.residual(z.mu)
res.mu <- mu - z.mu
res.mu <- attrcp(mu,res.mu)
class(res.mu) <- class(mu)
eof.res.mu <- EOF(res.mu)
plot(eof.res.mu)
```

For comparison the original data look like this:
```{r EOF-for-mu, warning=FALSE, message=FALSE}
## The residual of the wet-day precipitation
plot(EOF(mu))
```

```{r residual-for-fw, warning=FALSE, message=FALSE}
## Repeat with the wet-day frequency for the benefit of comparison
z.fw <- as.field(ds.fw)
#res.fw <- as.residual(ds.fw)
## Another way to estimate the residuals:
res.fw <- fw - z.fw
res.fw <- attrcp(fw,res.fw)
class(res.fw) <- class(fw)
eof.res.fw <- EOF(res.fw)
plot(eof.res.fw)
```

For comparison the original data look like this:
```{r EOF-for-fw, warning=FALSE, message=FALSE}
## The residual of the wet-day precipitation
plot(EOF(fw))
```

>>>>

```{r precipitation-intensity, warning=FALSE, message=FALSE}
## Show the mean seasonal cycle in the wet-day mean precipitation
plot(MU)
grid()
```

@@

```{r wet-day frequency, warning=FALSE, message=FALSE}
## Show the mean seasonal cycle in the wet-day frequency
plot(FW)
grid()
```

@@

```{r sesonal-precipitation, warning=FALSE, message=FALSE}
## Show the mean seasonal cycle in the total precipitation
MP <- 30*MP
attr(MP,'unit') <- 'mm/month'
plot(MP)
grid()
```

@@

```{r winter-rain, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## Estimate the PCA for the annual cycle to identify the general features:
pca.MU <- PCA(MU,n=10)
plot(pca.MU)
```

@@

```{r EOFs-in-precip-mu, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## The area mean wet-day mean precipitation
eof.mu <- EOF(mu)
## The area mean wet-day frequency
plot(eof.mu)

```{r EOFs-in-precip-fw, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## Show maps of trends for the wet-day mean precipitation
eof.fw <- EOF(fw)
## Map of trends in wet-day frequency:
plot(eof.fw)
```

```{r cca, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## The wet-day frequency and SLP
cca.fw <- CCA(eof.fw,eof.slp)
plot(cca.fw,new=FALSE)
cca.mu <- CCA(eof.mu,eof.es)
plot(cca.mu,new=FALSE)
```

```{r DS, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## The wet-day frequency and SLP
ds.fw <- DS(subset(eof.fw,ip=1:4),eof.slp,eofs=1:5)
plot(ds.fw,new=FALSE)
ds.mu <- DS(subset(eof.mu,ip=1:4),eof.es,eofs=1:7)
plot(ds.mu,new=FALSE)
```

Extract the residual and examine to see if there is any temporal-stpatial structure left: a tell-tale sign for noise is a flat eigenspectrum.

```{r residual-for-mu, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## The residual of the wet-day precipitation
z.mu <- as.field(ds.mu)
#res.mu <- as.residual(z.mu)
res.mu <- mu - z.mu
res.mu <- attrcp(mu,res.mu)
class(res.mu) <- class(mu)
eof.res.mu <- EOF(res.mu)
plot(eof.res.mu)
```

The residuals from the downscaling of $\mu$ still give eigenvalues which indicate that there are differnt coherent modes present associated with different variance. The spatial pattern is similar to the spatial patterns in the original data. This exercise suggests that the anaysis failed to capture any strong link between the predictor (vapour pressure estimated over the North Atlantic) and the predictand. One reason may be a shifting source region for moisture at different times of the year and from year to year.

For comparison the original data look like this:
```{r EOF-for-mu, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## The original of the wet-day precipitation
plot(EOF(mu))
```

```{r residual-for-fw, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## Repeat with the wet-day frequency for the benefit of comparison
z.fw <- as.field(ds.fw)
#res.fw <- as.residual(ds.fw)
## Another way to estimate the residuals:
res.fw <- fw - z.fw
res.fw <- attrcp(fw,res.fw)
class(res.fw) <- class(fw)
eof.res.fw <- EOF(res.fw)
plot(eof.res.fw)
```

The residual for the wet-day frequency ...

For comparison the original data look like this:
```{r EOF-for-fw, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## The original of the wet-day precipitation
plot(EOF(fw))
```

@@

```{r}
plot(gloss)
grid()
```

@@

```{r}
plot(newlyn)
grid()
```

@@

```{r sonel, warning=FALSE, message=FALSE}
sonel <- aggregate(station.sonel(),year,FUN='max',na.rm=TRUE)
plot(sonel)
grid()
```

@@

```{r trend-map, warning=FALSE, message=FALSE}
## Fudge it - pick some windstations, but do not use their data. Reset the
## coordinates to select a triangle of SLP from the EOBS data.
ss <- select.station(param='slp',src='ecad',cntr=c('France','UNITED KINGDOM'))
fx <- station(ss)
triangle <- subset(fx,is=1:3)
## The stations are not situated quite where there is valid slp data in the EOBS dataset
## It's only their coordinates that is used for interpolating SLP for three stations
## Using station objects as is-object makes the results a station object
attr(triangle,'longitude') <- c(-4,-2,-1.25)
attr(triangle,'latitude') <- c(50.50,51,49)
pp.eobs <- regrid(slp.eobs,is=triangle)
## Use triagulation for stations to estimate the wind
uv <- geostrophicwind(pp.eobs)
plot(annual(uv,nmin=300),main='zonal and meridional wind components',
     sub='Estimated from triangulation of mean sea-level pressure from EOBS',zoom=7)
grid()
```

@@

```{r trend-event-2006, warning=FALSE, message=FALSE}
plot(annual(uv,nmin=300),main='zonal and meridional wind components',
     xlim=c(2000,2010),map.show=FALSE)
grid()
```

@@

```{r zonal-wind-check1, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## Trend in upper wind direction
uam <- annual(subset(uv,is=2),nmin=300)
ds.u1 <- DS(uam,eof.slp,eofs=1:20)
plot(ds.u1)
```

@@

```{r zonal-wind-check2, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## Trend in upper wind direction
ds.u2 <- DS(subset(uam,it=c(1950,2005)),eof.slp,eofs=1:20)
plot(ds.u2)
```

@@

```{r reanalysis-wind-speed, warning=FALSE, message=FALSE}
## Estimate grostrophich wind
UV <- geostrophicwind(SLP,progressbar=FALSE)
```

@@
